---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r}
tf <- unzip("./activity.zip")
activity_data <- read.csv(tf)
activity_data$date<-as.Date(activity_data$date)
```


## What is mean total number of steps taken per day?
```{r histogram, message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
total_steps_per_day <- activity_data %>% 
                        group_by(date) %>% 
                        summarise(number_steps = sum(steps, na.rm = TRUE))
hist(total_steps_per_day$number_steps, xlab = "number of steps", ylim = c(0,35),
     main = "Total number of steps taken each day")
```

```{r mean mediam}
mean_steps_per_day <- mean(total_steps_per_day$number_steps)
median_steps_per_day <- median(total_steps_per_day$number_steps)
```

The mean of steps taken per day is 
`r sprintf("%.2f", mean_steps_per_day)`. The median of 
steps taken perday is `r sprintf("%.2f", median_steps_per_day)`.


## What is the average daily activity pattern?
```{r time series intervals}
average_steps_by_interval <- activity_data %>% 
                                group_by(interval) %>%
                                summarise(average_steps_by_interval = 
                                            mean(steps, na.rm = TRUE))
g <- ggplot(average_steps_by_interval, 
            aes(x=interval, y=average_steps_by_interval))
g + geom_line() + labs(title ="Average steps by interval", x = "Interval", 
                       y = "Average steps")

interval_max_average <- average_steps_by_interval$interval[which.max(
    average_steps_by_interval$average_steps_by_interval)]
```

The interval that contains the max average number of steps is 
`r interval_max_average`.

## Imputing missing values
```{r total missing values}
total_na <- sum(is.na(activity_data$steps))
```

The total number of missing values in the database is `r total_na`. 

```{r filling missing values}

ind_missing_intervals <- which(is.na(activity_data$steps))
#Replace the missing values with rounded step averages for matching interval on 
#other days
interpolated_steps <- replace(activity_data$steps, ind_missing_intervals, floor(average_steps_by_interval$average_steps_by_interval))

activity_data_interpolated <- data.frame(steps = interpolated_steps, 
                                         date = activity_data$date, 
                                         interval = activity_data$interval)
#Create histogram
total_steps_per_day_interpolated <- activity_data_interpolated %>% 
                        group_by(date) %>% 
                        summarise(number_steps = sum(steps, na.rm = TRUE))

hist(total_steps_per_day_interpolated$number_steps, xlab = "number of steps", 
     ylim = c(0,35), 
     main = "Total number of steps taken each day (filling missing values)")
```

```{r, mean and median}
mean_steps_per_day_interpolated <- mean(total_steps_per_day_interpolated$number_steps)
median_steps_per_day_interpolated <- median(total_steps_per_day_interpolated$number_steps)
```
The strategy used to fill the missing values was to replace the missing values 
with the rounded number of average steps of matching intervals of other days 
that data was collected.

The average steps taken per day after filling the missing values is 
`r sprintf("%.2f", mean_steps_per_day)`, and the median of steps taken per day after filling the 
missing values `r median_steps_per_day`. The average steps increased by 
`r  sprintf("%.2f", ((mean_steps_per_day_interpolated/mean_steps_per_day)-1)*100)`%, and 
the median by 
`r sprintf("%.2f", ((median_steps_per_day_interpolated/median_steps_per_day)-1)*100)`%. 
In terms of impact, the estimated steps created an extra 
`r sprintf("%i", sum(total_steps_per_day_interpolated$number_steps)-sum(total_steps_per_day$number_steps))`
steps in the elapsed `r nrow(total_steps_per_day)` days of data.

## Are there differences in activity patterns between weekdays and weekends?
```{r}
weekends <- weekdays(activity_data_interpolated$date) %in% 
    c("Saturday", "Sunday")
activity_data_interpolated$type_day <- factor(
    ifelse(weekends, "weekday", "weekend"))
# Calculate the average number of steps taken, averaged across all weekday days
# and weekend days
average_steps_by_interval <- activity_data_interpolated %>% 
                                     group_by(interval,type_day) %>%
                                     summarise(average_steps_by_interval = 
                                     mean(steps, na.rm = TRUE))
# Plot time series for weekdays and weekends
library(lattice)
xyplot(average_steps_by_interval ~ interval | type_day, 
       data = average_steps_by_interval, layout = c(1, 2), type = "l", 
       ylab = "number of steps")

```
