---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r}
tf <- unzip("./activity.zip")
activity_data <- read.csv(tf)
activity_data$date<-as.Date(activity_data$date)
```


## What is mean total number of steps taken per day?
```{r, message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
total_steps_per_day <- activity_data %>% 
                        group_by(date) %>% 
                        summarise(number_steps = sum(steps, na.rm = TRUE))
qplot(number_steps, data = total_steps_per_day, geom="histogram", binwidth=5000)
hist(total_steps_per_day$number_steps, xlab = "number of steps")
```

```{r, echo=TRUE}
mean_steps_per_day <- mean(total_steps_per_day$number_steps)
median_steps_per_day <- median(total_steps_per_day$number_steps)
```

The mean of steps taken per day is `r mean_steps_per_day`. The median of steps 
taken perday is `r median_steps_per_day`.


## What is the average daily activity pattern?
```{r time series intervals}
average_steps_by_interval <- activity_data %>% 
                                group_by(interval) %>%
                                summarise(average_steps_by_interval = 
                                            mean(steps, na.rm = TRUE))
g <- ggplot(average_steps_by_interval, 
            aes(x=interval, y=average_steps_by_interval))
g + geom_line()

#which.max(average_steps_by_interval$average_steps_by_interval)
interval_max_average <- average_steps_by_interval$interval[which.max(
    average_steps_by_interval$average_steps_by_interval)]

```

The interval that contains the max average number of steps is 
`r interval_max_average`.

## Imputing missing values
# Calculate and report the total number of missing values in the dataset
```{r}
total_na <- sum(is.na(activity_data$steps))
```

The total number of missing values in the database is `r total_na`. 

```{r filling missing values}
#missing_intervals <- activity_data$interval[is.na(activity_data$steps)]
ind_missing_intervals <- which(is.na(activity_data$steps))
#activity_data$steps[activity_data %in% ]
#ind <- match(missing_intervals, average_steps_by_interval$interval)
#Replace the missing values with rounded step averages for matching interval on 
#other days
interpolated_steps <- replace(activity_data$steps, ind_missing_intervals, floor(average_steps_by_interval$average_steps_by_interval))

activity_data_interpolated <- data.frame(steps = interpolated_steps, 
                                         date = activity_data$date, 
                                         interval = activity_data$interval)
#Create histogram
total_steps_per_day_interpolated <- activity_data_interpolated %>% 
                        group_by(date) %>% 
                        summarise(number_steps = sum(steps, na.rm = TRUE))

hist(total_steps_per_day_interpolated$number_steps, xlab = "number of steps")
```

```{r, mean and median}
mean_steps_per_day_interpolated <- mean(total_steps_per_day_interpolated$number_steps)
median_steps_per_day_interpolated <- median(total_steps_per_day_interpolated$number_steps)
```

The average steps taken per day after filling the missing values is 
`r mean_steps_per_day`, and the median of steps taken per day after filling the 
missing values `r median_steps_per_day`. The average steps increased by 
`r  ((mean_steps_per_day_interpolated/mean_steps_per_day)-1)*100`% and 
the median by 
`r ((median_steps_per_day_interpolated/median_steps_per_day)-1)*100`% In 
terms of impact, the estimated steps created an extra 
`r sum(total_steps_per_day_interpolated$number_steps)-sum(total_steps_per_day$number_steps)`
steps in the elapsed `r nrow(total_steps_per_day)` days of data.

## Are there differences in activity patterns between weekdays and weekends?
```{r}
weekends <- weekdays(activity_data_interpolated$date) %in% 
    c("Saturday", "Sunday")
activity_data_interpolated$type_day <- factor(
    ifelse(weekends, "weekday", "weekend"))

library(lattice)
average_steps_by_interval <- activity_data_interpolated %>% 
                                     group_by(interval,type_day) %>%
                                     summarise(average_steps_by_interval = 
                                     mean(steps, na.rm = TRUE))

xyplot(average_steps_by_interval ~ interval | type_day, 
       data = average_steps_by_interval, layout = c(1, 2), type = "l", 
       ylab = "number of steps")

```
